MCC Models

### 2019-03-10 papers/full_paper/models.Rmd
parameter free model predictions

library(tidyverse)
library(knitr)
library(ggthemes)
library(langcog)
library(rwebppl)
library(coda)

@@

```{r rsaUtils, include = FALSE}
rsaUtils <- '
var all_objects = [
{ shape: "triangle", id:1, location: 1},  
{ shape: "triangle", id:2, location: 2},
{ shape: "circle", id:1, location: 2}
]

@@

## RSA model code

```{r rsaModel}
rsaModel <- '
var pragmaticListener = function(utterance, priorProbs){
Infer({method: "enumerate", model: function(){
var lexiconName = sample(LexiconPrior);
var obj = sample( Categorical({vs: all_objects, ps: priorProbs}));
var S1 = speaker(obj, lexiconName, priorProbs);
observe(S1, utterance)
return obj.shape == "circle" ? 1 : 0
}})
}

var speaker = function(obj, lexiconName, priorProbs){
Infer({method: "enumerate", model: function(){
var utterance = utterancePrior(obj, lexiconName);
var L0 = literalListener(utterance, priorProbs);
 factor(speakerOptimality * L0.score(obj.shape))
return utterance
}})
}

var literalListener = function(utterance, priorProbs){
Infer({method: "enumerate", model: function(){
var lexiconName = sample(LexiconPrior); 
var lexicon = lexiconObject[lexiconName];
var obj = sample( Categorical({vs: all_objects, ps: priorProbs}));
if ("label" in utterance) {
 var truthValue = lexicon(utterance, obj);
 condition(truthValue)
}
if (utterance.point) {
 var truthValuePoint = point(utterance, obj);
 condition(truthValuePoint)
}
return obj.shape 
}})
}
'
```

@@

```{r ex3 data}
# load data from experiment 3
adult_ex3_data <-read_csv(file="../../stats/data/adult_ex3.csv") 
```

```{r}
# load data from prior manipulation experiments to set priors
adult_ex2_novelty_data <- read_csv(file="../../stats/data/adult_ex2_novelty.csv")
adult_ex2_preference_data <- read_csv(file="../../stats/data/adult_ex2_preference.csv")
```

# create datafiles with model parameters
predPriorEx3 <- data.frame(
  common_ground_manipulation = c("preference","preference","preference","preference","novelty","novelty","novelty","novelty"),
  speaker = c("same_speaker","same_speaker","different_speaker","different_speaker","same_speaker","same_speaker","different_speaker","different_speaker"),
  alignment = c("congruent","incongruent","congruent","incongruent","congruent","incongruent","congruent","incongruent"))

@@

predPriorEx3$prior = list(
    c(1-prior_pref_same$prior,1-prior_pref_same$prior,prior_pref_same$prior), 
    c(prior_pref_same$prior,prior_pref_same$prior,1-prior_pref_same$prior), 
    c(1-prior_pref_diff$prior,1-prior_pref_diff$prior,prior_pref_diff$prior), 
    c(prior_pref_diff$prior,prior_pref_diff$prior,1-prior_pref_diff$prior),
    c(1-prior_novel_same$prior,1-prior_novel_same$prior,prior_novel_same$prior), 
    c(prior_novel_same$prior,prior_novel_same$prior,1-prior_novel_same$prior), 
    c(1-prior_novel_diff$prior,1-prior_novel_diff$prior,prior_novel_diff$prior), 
    c(prior_novel_diff$prior,prior_novel_diff$prior,1-prior_novel_diff$prior))

# data frame to keep track of unique conditions
allConditionsEx3 <- predPriorEx3 %>%
  select(-prior)

@@

```{r model rsa noise ex3, include = FALSE}
predictions.rsa.ex3 <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.predPrior

var model = function(){
   var noise = uniformDrift({a: 0, b:1, width: 0.1})

  var conditionOutput = map(function(conditionInfo){

  var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var conditionSpecificPriors =  _.filter(allPriorProbs, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var modelPredictions = pragmaticListener({label: "dax", point: 2 }, conditionSpecificPriors[0].prior)
   var noisyModelPredictions = addNoise(modelPredictions, noise)

   map(function(d){ 
      observe(noisyModelPredictions, d.correct_inf)
    }, conditionSpecificData)

  return  [conditionInfo.common_ground_manipulation + "/" + conditionInfo.speaker + "/" +  conditionInfo.alignment, Math.exp(noisyModelPredictions.score(1))]

}, allConditions)

  return extend(_.fromPairs(conditionOutput), {noise: noise})

```{r predictions pragmatic model ex 3, include = FALSE}

# ex3_pragm_model <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.rsa.ex3, sep='\n'),
#   data = list(allData = adult_ex3_data, allConditions = allConditionsEx3, predPrior = predPriorEx3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 2000, burn = 500, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex3_pragm_model.rds")

ex3_pragm_model <- readRDS("../../stats/saves/ex3_pragm_model.rds")

# model predictions
ex3_pragm_model_pred <- ex3_pragm_model %>%
  filter(!(Parameter %in% c("noise")))  %>%
  separate(Parameter, into = c("common_ground_manipulation","speaker", "alignment"), sep="/")%>%
  mutate(model="pragmatic")

## noise parameter
ex3_pragm_model_noise <- ex3_pragm_model %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

@@

```{r model prior only ex3 , include = FALSE}
predictions.prior.only.ex3 <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.predPrior

var model = function(){
   var noise = uniformDrift({a: 0, b:1, width: 0.1})

  var conditionOutput = map(function(conditionInfo){

  var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var conditionSpecificPriors =  _.filter(allPriorProbs, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

  var modelPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: conditionSpecificPriors[0].prior}));
      return obj.shape == "circle" ? 1 : 0
  }})

   var noisyModelPredictions = addNoise(modelPredictions, noise)

   map(function(d){ 
      observe(noisyModelPredictions, d.correct_inf)
    }, conditionSpecificData)

  return  [conditionInfo.common_ground_manipulation + "/" + conditionInfo.speaker + "/" +  conditionInfo.alignment, Math.exp(noisyModelPredictions.score(1))]

}, allConditions)

return extend(_.fromPairs(conditionOutput), {noise: noise})

```{r predictions prior only model ex3, include = FALSE}

# ex3_prior_only_model <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.prior.only.ex3, sep='\n'),
#   data = list(allData = adult_ex3_data, allConditions = allConditionsEx3, predPrior = predPriorEx3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 2000, burn = 500, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex3_prior_only_model.rds")

ex3_prior_only_model <- readRDS("../../stats/saves/ex3_prior_only_model.rds")

# model predictions
ex3_prior_only_model_pred <- ex3_prior_only_model %>%
  filter(!(Parameter %in% c("noise")))  %>%
  separate(Parameter, into = c("common_ground_manipulation", "speaker", "alignment"), sep="/")%>%
  mutate(model="prior_only")

## noise parameter
ex3_prior_only_model_noise <- ex3_prior_only_model %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

@@

```{r model flat prior ex3, include = FALSE}
predictions.flat.prior.ex3 <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.predPrior

var model = function(){
   var noise = uniformDrift({a: 0, b:1, width: 0.1})

  var conditionOutput = map(function(conditionInfo){

  var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var conditionSpecificPriors = [.5, .5, .5]

   var modelPredictions = pragmaticListener({label: "dax", point: 2 }, conditionSpecificPriors)
   var noisyModelPredictions = addNoise(modelPredictions, noise)

   map(function(d){ 
      observe(noisyModelPredictions, d.correct_inf)
    }, conditionSpecificData)

  return  [conditionInfo.common_ground_manipulation + "/" + conditionInfo.speaker + "/" +  conditionInfo.alignment, Math.exp(noisyModelPredictions.score(1))]

}, allConditions)

  return extend(_.fromPairs(conditionOutput), {noise: noise})

```{r predictions prior only model ex3, include = FALSE}

# ex3_flat_prior_model <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.flat.prior.ex3, sep='\n'),
#   data = list(allData = adult_ex3_data, allConditions = allConditionsEx3, predPrior = predPriorEx3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 2000, burn = 500, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex3_flat_prior_model.rds")

ex3_flat_prior_model <- readRDS("../../stats/saves/ex3_flat_prior_model.rds")

# model predictions
ex3_flat_prior_model_pred <- ex3_flat_prior_model %>%
  filter(!(Parameter %in% c("noise")))  %>%
  separate(Parameter, into = c("common_ground_manipulation", "speaker", "alignment"), sep="/")%>%
  mutate(model="flat_prior")

## noise parameter
ex3_flat_prior_model_noise <- ex3_flat_prior_model %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

@@

# summarize model predictions
ex_3_model_pred <- bind_rows(
  ex3_pragm_model_pred,
  ex3_prior_only_model_pred,
  ex3_flat_prior_model_pred
)%>%
  group_by(model,common_ground_manipulation,speaker,alignment) %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value), 
            ci_upper = hdi_upper(value))

@@

```{r plot ex3 model predictions and data}
ggplot(data = ex_3_pred, aes(x = alignment, y = mean, col = model)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), size = .4,  position = position_dodge(width = .5), pch = 5)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Choosing More Informative")+
  facet_grid (~ common_ground_manipulation + speaker) +
  theme_few() + 
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  scale_color_manual(name="Model",
                     breaks=c("data","pragmatic","flat_prior","prior_only"),
                     values= c("black","#859900","#6c71c4","#b58900"))+
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))
```

@@

```{r plot }

@@

### Model comparison

```{r model likelihood rsa ex3, include = FALSE}
loglike.rsa.ex3 <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.predPrior

var model = function(){
   var noise = uniformDrift({a: 0, b:1, width: 0.1})

  var conditionOutput = map(function(conditionInfo){

  var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var conditionSpecificPriors =  _.filter(allPriorProbs, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var modelPredictions = pragmaticListener({label: "dax", point: 2 }, conditionSpecificPriors[0].prior)

    var noisyModelPredictions = addNoise(modelPredictions, noise)

    var loglike = Binomial({
        p: Math.exp(noisyModelPredictions.score(1)), 
        n: conditionSpecificData.length}).score(
      sum(_.map(conditionSpecificData, "correct_inf"))
    )

  return  loglike

}, allConditions)

 return sum(conditionOutput)

```{r model likelihood prior only ex3 , include = FALSE}
loglike.prior.only.ex3 <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.predPrior

var model = function(){
   var noise = uniformDrift({a: 0, b:1, width: 0.1})

  var conditionOutput = map(function(conditionInfo){

  var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var conditionSpecificPriors =  _.filter(allPriorProbs, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

  var modelPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: conditionSpecificPriors[0].prior}));
      return obj.shape == "circle" ? 1 : 0
  }})

   var noisyModelPredictions = addNoise(modelPredictions, noise)

    var loglike = Binomial({
        p: Math.exp(noisyModelPredictions.score(1)), 
        n: conditionSpecificData.length}).score(
      sum(_.map(conditionSpecificData, "correct_inf"))
    )

  return  loglike

}, allConditions)

 return sum(conditionOutput)

```{r model likelihood flat prior ex3, include = FALSE}
loglike.flat.prior.ex3 <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.predPrior

var model = function(){
   var noise = uniformDrift({a: 0, b:1, width: 0.1})

  var conditionOutput = map(function(conditionInfo){

  var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var conditionSpecificPriors = [.5, .5, .5]

   var modelPredictions = pragmaticListener({label: "dax", point: 2 }, conditionSpecificPriors)
   var noisyModelPredictions = addNoise(modelPredictions, noise)

    var loglike = Binomial({
        p: Math.exp(noisyModelPredictions.score(1)), 
        n: conditionSpecificData.length}).score(
      sum(_.map(conditionSpecificData, "correct_inf"))
    )

  return  loglike

}, allConditions)

 return sum(conditionOutput)

```{r ex3 model comparison}
# ex3_pragm_model_loglike <- webppl(
#   program_code = paste(rsaUtils, rsaModel, loglike.rsa.ex3, sep='\n'),
#   data = list(allData = adult_ex3_data, allConditions = allConditionsEx3, predPrior = predPriorEx3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4, 
#   cores = 4,
#   inference_opts = list(method = "forward", samples = 5000)
# ) %>%
#   saveRDS("../../stats/saves/ex3_pragm_model_loglike.rds")

ex3_pragm_model_loglike <- readRDS("../../stats/saves/ex3_pragm_model_loglike.rds")%>%
  mutate(model = "pragmatic")%>%
  group_by(model)%>%
  summarize(logP = logSumExp(value))

# ex3_prior_only_model_loglike <- webppl(
#   program_code = paste(rsaUtils, rsaModel, loglike.prior.only.ex3, sep='\n'),
#   data = list(allData = adult_ex3_data, allConditions = allConditionsEx3, predPrior = predPriorEx3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4, 
#   cores = 4,
#   inference_opts = list(method = "forward", samples = 5000)
# )%>%
#   saveRDS("../../stats/saves/ex3_prior_only_model_loglike.rds")

ex3_prior_only_model_loglike <- readRDS("../../stats/saves/ex3_prior_only_model_loglike.rds")%>%
  mutate(model = "prior_only")%>%
  group_by(model)%>%
  summarize(logP = logSumExp(value))

# ex3_flat_prior_model_loglike <- webppl(
#   program_code = paste(rsaUtils, rsaModel, loglike.flat.prior.ex3, sep='\n'),
#   data = list(allData = adult_ex3_data, allConditions = allConditionsEx3, predPrior = predPriorEx3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4, 
#   cores = 4,
#   inference_opts = list(method = "forward", samples = 5000)
# )%>%
#   saveRDS("../../stats/saves/ex3_flat_prior_model_loglike.rds")

ex3_flat_prior_model_loglike <- readRDS("../../stats/saves/ex3_flat_prior_model_loglike.rds")%>%
  mutate(model = "flat_prior")%>%
  group_by(model)%>%
  summarize(logP = logSumExp(value))

ex3_model_comparison <- bind_rows(
  ex3_pragm_model_loglike,
  ex3_prior_only_model_loglike,
  ex3_flat_prior_model_loglike
)%>%
  mutate(model = reorder(model,logP))

#saveRDS(ex3_model_comparison, "../../stats/saves/ex3_model_comparison.rds")

```
##### Plot for likelihoods

```{r}
ggplot(ex3_model_comparison, aes(x = model, y = logP, col = model))+
  geom_segment( aes(x=model, xend=model, y=0, yend=logP), col = "black", size = 1)+
  geom_point(size = 7)+
  theme_few()+ 
  xlab("Model")+
  ylab("Log-likelihood")+
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))+
  guides(col = F)+
  scale_color_manual(breaks=c("pragmatic","prior_only","flat_prior"),
                    values= c("#859900","#b58900","#6c71c4"))
```

##### Bayes factors

```{r}
ex3_bf <- data.frame(
  comparison = c("pragmatic vs. prior_only","pragmatic vs. flat_prior", "flat_rior vs. prior_only"),
  bayes_factor = c(
    exp(ex3_model_comparison$logP[ex3_model_comparison$model== "pragmatic"] - ex3_model_comparison$logP[ex3_model_comparison$model== "prior_only"]),
    exp(ex3_model_comparison$logP[ex3_model_comparison$model== "pragmatic"] - ex3_model_comparison$logP[ex3_model_comparison$model== "flat_prior"]),
    exp(ex3_model_comparison$logP[ex3_model_comparison$model== "prior_only"] - ex3_model_comparison$logP[ex3_model_comparison$model== "flat_prior"]))
)

@@

### Model parameters and data
```{r ex4 data}
# load data  from experiment 4 
adult_ex4_data <-read_csv(file="../../stats/data/adult_ex4.csv") 
```

```{r}
# load data from prior strength manipulation experiments to set priors
adult_ex4_prior_data <-read_csv(file="../../stats/data/adult_ex4_prior.csv") 
```

@@

# combining all priors into one data frame
predPriorEx4 = rbind(strongPrior,mediumPrior,weakPrior)

# data frame to keep track of unique conditions
allConditionsEx4 <- predPriorEx4 %>%
  select(-prior)

@@

```{r model rsa noise ex4, include = FALSE}
predictions.rsa.ex4 <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.predPrior
var noise = dataFromR.noise[0]

var conditionOutput = map(function(conditionInfo){

    var conditionSpecificPriors =  _.filter(allPriorProbs, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment, prior_manipulation : conditionInfo.prior_manipulation})

    var modelPredictions = pragmaticListener({label: "dax", point: 2 }, conditionSpecificPriors[0].prior)

    var noisyModelPredictions = addNoise(modelPredictions, noise)

    return  extend([conditionInfo.common_ground_manipulation + "/" + conditionInfo.speaker + "/" +  conditionInfo.alignment+ "/" + conditionInfo.prior_manipulation, Math.exp(noisyModelPredictions.score(1))])

  }, allConditions)

   conditionOutput

@@

```{r predictions pragmatic model, include = FALSE}

ex4_pragm_model <- webppl(
  program_code = paste(rsaUtils, rsaModel, predictions.rsa.ex4, sep='\n'),
  data = list(allData = adult_ex4_data, allConditions = allConditionsEx4, predPrior = predPriorEx4, noise = ex3_pragm_model_noise$mean),
  data_var = "dataFromR"
)

# model predictions
ex4_pragm_model_pred <- ex4_pragm_model %>%
  separate(`0`, into = c("common_ground_manipulation", "speaker", "alignment", "prior_manipulation"), sep="/")%>%
  mutate(value = `1`,
         model="pragmatic",
         prior_manipulation = relevel(as.factor(prior_manipulation), ref = "weak"))%>%
  select(-`1`)

@@

```{r model prior only ex4 , include = FALSE}
predictions.prior.only.ex4 <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.predPrior
var noise = dataFromR.noise[0]

var conditionOutput = map(function(conditionInfo){

   var conditionSpecificPriors =  _.filter(allPriorProbs, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment, prior_manipulation : conditionInfo.prior_manipulation})

  var modelPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: conditionSpecificPriors[0].prior}));
      return obj.shape == "circle" ? 1 : 0
  }})

   var noisyModelPredictions = addNoise(modelPredictions, noise)

  return  extend([conditionInfo.common_ground_manipulation + "/" + conditionInfo.speaker + "/" +  conditionInfo.alignment+ "/" + conditionInfo.prior_manipulation, Math.exp(noisyModelPredictions.score(1))])

}, allConditions)

conditionOutput

@@

### Model comparison

>>>>

library(tidyverse)
library(knitr)
library(ggthemes)
library(langcog)
library(rwebppl)
library(coda)
library(ggpubr)

@@

```{r rsaUtils}
rsaUtils <- '
var all_objects = [
{ shape: "triangle", id:1, location: 1},  
{ shape: "triangle", id:2, location: 2},
{ shape: "circle", id:1, location: 2}
]

@@

## RSA model code
```{r rsa model}
rsaModel <- '
var literalListener = cache(function(utterance, priorProbs){
Infer({method: "enumerate", model: function(){
var lexiconName = sample(LexiconPrior); 
var lexicon = lexiconObject[lexiconName];
var obj = sample( Categorical({vs: all_objects, ps: priorProbs}));
if ("label" in utterance) {
 var truthValue = lexicon(utterance, obj);
 condition(truthValue)
}
if (utterance.point) {
 var truthValuePoint = point(utterance, obj);
 condition(truthValuePoint)
}
return obj.shape 
}})
})

@@

```{r ex3 data adults}
# load data from experiment 3
adult_ex3_data <-read_csv(file="../../stats/data/adult_ex3.csv") 
```

```{r ex 1 and 2 data adults}
# adults exp 1 
adult_ex1_data <- read_csv(file="../../stats/data/adult_ex1.csv")
# load data from prior manipulation experiments to set priors
adult_ex2_novelty_data <- read_csv(file="../../stats/data/adult_ex2_novelty.csv")
adult_ex2_preference_data <- read_csv(file="../../stats/data/adult_ex2_preference.csv")
```

#### Inferring speaker optimality parameter and setting priors

```{r model parameter adults}
speakOptParam <- '
var data = dataFromR.data

  var speakerOptimality = uniformDrift({a: -3, b: 7, width: 0.2})

  var priorProbs = [.5, .5, .5]

  var rsaPredictions = pragmaticListener({label: "dax", point: 2 },
   priorProbs, speakerOptimality) 
   observe(rsaPredictions, row.correct)

  }, data)

  return speakerOptimality

```{r inferring speaker optimality parameter adults}
# adult_model_param<- webppl(
#   program_code = paste(rsaUtils, rsaModel, speakOptParam , sep='\n'),
#   data = list(data = adult_ex1_data),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# )%>%
#   saveRDS("../../stats/saves/adult_model_param.rds")

adult_model_param <- readRDS("../../stats/saves/adult_model_param.rds")%>%
  select(value) %>%
  summarise(so_map = estimate_mode(value),
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

```{r visualize speaker optimality parameter adults}
# visualize speaker optimality parameter
plot_adult_model_speak_opt <- readRDS("../../stats/saves/adult_model_param.rds")%>%
  select(value)

ggplot(plot_adult_model_speak_opt, aes(value))+
  geom_density()+
  xlab("Speaker optimality")+
  theme_few()
```

```{r model setup ex 3 adults, include = FALSE}
# create datafiles with model parameters
model_params_ex3 <- data.frame(
  common_ground_manipulation = c("preference","preference","preference","preference","novelty","novelty","novelty","novelty"),
  speaker = c("same_speaker","same_speaker","different_speaker","different_speaker","same_speaker","same_speaker","different_speaker","different_speaker"),
  alignment = c("congruent","incongruent","congruent","incongruent","congruent","incongruent","congruent","incongruent"),
  speak_opt = adult_model_param$so_map)

@@

model_params_ex3$prior = list(
    c(1-prior_pref_same$prior,1-prior_pref_same$prior,prior_pref_same$prior), 
    c(prior_pref_same$prior,prior_pref_same$prior,1-prior_pref_same$prior), 
    c(1-prior_pref_diff$prior,1-prior_pref_diff$prior,prior_pref_diff$prior), 
    c(prior_pref_diff$prior,prior_pref_diff$prior,1-prior_pref_diff$prior),
    c(1-prior_novel_same$prior,1-prior_novel_same$prior,prior_novel_same$prior), 
    c(prior_novel_same$prior,prior_novel_same$prior,1-prior_novel_same$prior), 
    c(1-prior_novel_diff$prior,1-prior_novel_diff$prior,prior_novel_diff$prior), 
    c(prior_novel_diff$prior,prior_novel_diff$prior,1-prior_novel_diff$prior))

@@

```{r model rsa noise ex3 adults, include = FALSE}
predictions.rsa.ex3 <-'

var allData = dataFromR.data

    var modelPredictions = pragmaticListener({label: "dax", point: 2 }, row.prior,row.speak_opt)

    return extend([row.common_ground_manipulation + "/" + row.speaker + "/" + row.alignment + "/" + row.correct_inf, Math.exp(modelPredictions.score(1))])

```{r predictions pragmatic model ex 3 adults, include = FALSE}
ex3_pragm_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, predictions.rsa.ex3 , sep='\n'),
  data =list(data = model_params_ex3),
  data_var = "dataFromR"
)

# model predictions
ex3_pragm_model_predictions <- ex3_pragm_model %>%
  separate(`0`, into = c("common_ground_manipulation", "speaker", "alignment"), sep="/")%>%
  mutate(value = `1`,
         model="pragmatic")%>%
  select(-`1`)

@@

```{r model prior only ex3 adults, include = FALSE}
predictions.prior.only.ex3 <-'

var allData = dataFromR.data

  var modelPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: row.prior}));
      return obj.shape == "circle" ? 1 : 0
  }})

    return extend([row.common_ground_manipulation + "/" + row.speaker + "/" + row.alignment + "/" + row.correct_inf, Math.exp(modelPredictions.score(1))])

```{r predictions prior only model ex3 adults, include = FALSE}
ex3_prior_only_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, predictions.prior.only.ex3 , sep='\n'),
  data =list(data = model_params_ex3),
  data_var = "dataFromR"
)

# model predictions
ex3_prior_only_model_predictions <- ex3_prior_only_model %>%
  separate(`0`, into = c("common_ground_manipulation", "speaker", "alignment"), sep="/")%>%
  mutate(value = `1`,
         model="prior_only")%>%
  select(-`1`)
```

@@

```{r model flat prior ex3 adults, include = FALSE}
predictions.flat.prior.ex3 <-'
var allData = dataFromR.data

    var modelPredictions = pragmaticListener({label: "dax", point: 2 }, prior,row.speak_opt)

    return extend([row.common_ground_manipulation + "/" + row.speaker + "/" + row.alignment + "/" + row.correct_inf, Math.exp(modelPredictions.score(1))])

output
'
```

```{r predictions prior only model ex3 adults, include = FALSE}
ex3_flat_prior_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, predictions.flat.prior.ex3 , sep='\n'),
  data =list(data = model_params_ex3),
  data_var = "dataFromR"
)

# model predictions
ex3_flat_prior_model_predictions <- ex3_flat_prior_model %>%
  separate(`0`, into = c("common_ground_manipulation", "speaker", "alignment"), sep="/")%>%
  mutate(value = `1`,
         model="flat_prior")%>%
  select(-`1`)

@@

# summarize model predictions
ex_3_model_pred <- bind_rows(
  ex3_pragm_model_predictions,
  ex3_prior_only_model_predictions,
  ex3_flat_prior_model_predictions
)%>%
  group_by(model,common_ground_manipulation,speaker,alignment)%>%
  summarise(mean = mean(value))%>%
  mutate(ci_lower = mean,
         ci_upper = mean)

@@

```{r plot ex3 model predictions and data adults}
ggplot(data = ex_3_pred, aes(x = alignment, y = mean, col = model)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), size = .4,  position = position_dodge(width = .5), pch = 5)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Choosing More Informative")+
  facet_grid (~ common_ground_manipulation + speaker) +
  theme_few() + 
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  scale_color_manual(name="Model",
                     breaks=c("data","pragmatic","flat_prior","prior_only"),
                     values= c("black","#859900","#6c71c4","#b58900"))+
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))
```

@@

```{r correlation plot ex3 adults}

@@

## Model comparison

```{r log likes ex3 adults}
# calculating model log-likelihoods
log_like_pragmatic_ex3 <- adult_ex3_data %>%
  left_join(ex3_pragm_model_predictions)%>%
  mutate(like = ifelse(correct_inf == "1",value,1-value),
         log_like = log(like))%>%
  group_by(model)%>%
  summarise(model_log_like = sum(log_like))

log_like_flat_prior_ex3 <- adult_ex3_data %>%
  left_join(ex3_flat_prior_model_predictions)%>%
  mutate(like = ifelse(correct_inf == "1",value,1-value),
         log_like = log(like))%>%
  group_by(model)%>%
  summarise(model_log_like = sum(log_like))

log_like_prior_only_ex3 <- adult_ex3_data %>%
  left_join(ex3_prior_only_model_predictions)%>%
  mutate(like = ifelse(correct_inf == "1",value,1-value),
         log_like = log(like))%>%
  group_by(model)%>%
  summarise(model_log_like = sum(log_like))

ex3_model_comparison <- bind_rows(
  log_like_pragmatic_ex3,
  log_like_flat_prior_ex3,
  log_like_prior_only_ex3
)%>%
  mutate(model = reorder(model,model_log_like))
```

### Plot for likelihoods

```{r likelihood plot adults}
ggplot(ex3_model_comparison, aes(x = model, y = model_log_like, col = model))+
  geom_segment( aes(x=model, xend=model, y=-600, yend=model_log_like), col = "black", size = 1)+
  geom_point(size = 7)+
  theme_few()+ 
  xlab("Model")+
  ylab("Log-likelihood")+
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))+
  guides(col = F)+
  scale_color_manual(breaks=c("pragmatic","prior_only","flat_prior"),
                    values= c("#b58900","#859900","#6c71c4"))
```

### Bayes factors

```{r bayes factors ex3 adults}
ex3_bf <- data.frame(
  comparison = c("pragmatic vs. flat_prior","pragmatic vs. prior_only", "flat_rior vs. prior_only"),
  log_bayes_factor = c(
    log_like_pragmatic_ex3$model_log_like - log_like_flat_prior_ex3$model_log_like,
    log_like_pragmatic_ex3$model_log_like- log_like_prior_only_ex3$model_log_like,
    log_like_flat_prior_ex3$model_log_like - log_like_prior_only_ex3$model_log_like)
)

@@

### Model parameters and data
```{r ex4 data adults}
# load data  from experiment 4 
adult_ex4_data <-read_csv(file="../../stats/data/adult_ex4.csv") 
```

```{r prior pre-tests data adults}
# load data from prior strength manipulation experiments to set priors
adult_ex4_prior_data <-read_csv(file="../../stats/data/adult_ex4_prior.csv") 
```

@@

# combining all priors into one data frame
model_params_ex4 = rbind(strongPrior,mediumPrior,weakPrior)%>%
  mutate(speak_opt = adult_model_param$so_map)

@@

```{r model rsa noise ex4 adults, include = FALSE}
predictions.rsa.ex4 <-'

var allData = dataFromR.data

    var modelPredictions = pragmaticListener({label: "dax", point: 2 }, row.prior,row.speak_opt)

    return extend([row.common_ground_manipulation + "/" + row.prior_manipulation + "/" + row.speaker + "/" + row.alignment + "/" + row.correct_inf, Math.exp(modelPredictions.score(1))])

@@

```{r predictions pragmatic model ex 4 adults, include = FALSE}
ex4_pragm_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, predictions.rsa.ex4 , sep='\n'),
  data =list(data = model_params_ex4),
  data_var = "dataFromR"
)

# model predictions
ex4_pragm_model_predictions <- ex4_pragm_model %>%
  separate(`0`, into = c("common_ground_manipulation","prior_manipulation", "speaker", "alignment"), sep="/")%>%
  mutate(value = `1`,
         model="pragmatic")%>%
  select(-`1`)

@@

```{r model prior only ex4 adults, include = FALSE}
predictions.prior.only.ex4 <-'

var allData = dataFromR.data

  var modelPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: row.prior}));
      return obj.shape == "circle" ? 1 : 0
  }})

    return extend([row.common_ground_manipulation + "/" + row.prior_manipulation + "/" + row.speaker + "/" + row.alignment + "/" + row.correct_inf, Math.exp(modelPredictions.score(1))])

@@

## Model comparison
